---
title: "R Notebook"
output: html_notebook
---






```{r}

library(tidyverse)
library(proDA)
library(DEP)
library(SummarizedExperiment)
library(cowplot)
source("run_dep.R")
source("run_proDA.R")
source("run_qprot.R")
source("run_perseus.R")
```


Load the data from the de Graaf et al. (2014) pubication that studied the effect of 
prostaglandin E2 (PGE2) on phosphorylation at different timepoints. The data was
uploaded to the ProteomeXchange with the identifier PXD000293. I have extracted
the proteinGroups MaxQuant output file. I will now load that file

```{r}

full_data <- read_tsv(
  "../data/degraaf_phosphorylation/degraaf_proteinGroups.txt",
  col_types = list(
    `LFQ intensity OR8_5min_C3` = "n",
    `LFQ intensity OR9_10min_B1` = "n",
    `LFQ intensity OR9_10min_A2` = "n",
    `LFQ intensity OR9_30min_B3` = "n",
    `LFQ intensity OR9_5min_B3` = "n",
    `LFQ intensity OR9_10min_A2` = "n",
    `LFQ intensity OR9_10min_A3` = "n"
  )
) 

tidy_data <- full_data %>%
  dplyr::select(ProteinID=`Protein IDs`, starts_with("LFQ Intensity ")) %>%
  gather(Sample, Intensity, starts_with("LFQ intensity ")) %>%
  mutate(MS_Run = str_match(Sample, "^LFQ intensity (OR\\d)_(\\d+)min_([A-C])(\\d)$")[,2]) %>% 
  mutate(Timepoint = as.numeric(str_match(Sample, "^LFQ intensity (OR\\d)_(\\d+)min_([A-C])(\\d)$")[,3])) %>% 
  mutate(BioReplicate = str_match(Sample, "^LFQ intensity (OR\\d)_(\\d+)min_([A-C])(\\d)$")[,4]) %>% 
  mutate(TechReplicate = as.numeric(str_match(Sample, "^LFQ intensity (OR\\d)_(\\d+)min_([A-C])(\\d)$")[,5])) %>% 
  mutate(SampleName = paste0("t", str_pad(Timepoint, width=2, pad="0"),
                             "_", BioReplicate, "_Rep", TechReplicate,  "_", MS_Run)) 

tidy_data

```




```{r}

data <- tidy_data %>%
  filter(Timepoint == 0) %>%
  mutate(Intensity = ifelse(Intensity == 0, NA, log2(Intensity))) %>%
  select(ProteinID, SampleName, Intensity) %>%
  spread(SampleName, Intensity) %>%
  column_to_rownames("ProteinID") %>%
  as.matrix() 

data <- data[rowSums(is.na(data)) < 18, ]

data_norm <- median_normalization(data)



annotation_df <- tidy_data %>%
  filter(Timepoint == 0) %>%
  select(SampleName, BioReplicate, TechReplicate, MS_Run) %>%
  distinct() %>%
  mutate(TechReplicate = as.factor(TechReplicate)) %>%
  arrange(BioReplicate, TechReplicate) %>%
  as.data.frame() %>%
  column_to_rownames("SampleName")


```



```{r}
data_for_plot <- data_norm
data_for_plot[is.na(data_for_plot)] <- 0
pheatmap::pheatmap(data_for_plot, show_rownames = FALSE, 
                   annotation_col = annotation_df)

```



```{r}
set.seed(1)
zero_parameters <- proDA(data_norm, col_data = annotation_df, verbose=TRUE)

```

```{r}
write_rds(zero_parameters, "tmp/de_graaf/zero_info_parameters.RDS")
zero_parameters <- read_rds("tmp/de_graaf/zero_info_parameters.RDS")
```



```{r}
sample_distances <- dist_approx(zero_parameters)$mean
protein_distances <- dist_approx( zero_parameters, by_sample = FALSE)$mean
```

```{r}
cowplot::plot_grid(pheatmap::pheatmap(data_norm, show_rownames = FALSE, 
                   cluster_rows = hclust(protein_distances), 
                   cluster_cols = hclust(sample_distances),
                   annotation_col = annotation_df, silent=TRUE)$gtable)


as.data.frame(cmdscale(sample_distances)) %>%
  rownames_to_column("Names") %>%
  inner_join(rownames_to_column(annotation_df, "Names")) %>%
  ggplot(aes(x = V1, y=V2, color=MS_Run, shape = BioReplicate)) +
    geom_point(size=4) +
    ggrepel::geom_text_repel(aes(label = Names), point.padding = 0.1, box.padding = 0.3, size=3) +
    coord_fixed()
```


```{r}
annotation_df <- tidy_data %>%
  filter(Timepoint == 0) %>%
  select(SampleName, BioReplicate, TechReplicate, MS_Run) %>%
  distinct() %>%
  mutate(TechReplicate = as.factor(TechReplicate)) %>%
  arrange(BioReplicate, TechReplicate) %>%
  as.data.frame() %>%
  column_to_rownames("SampleName")


sel_samples_3v3 <- list(Syn_Cond1 = paste0("t00_", c("A_Rep2", "C_Rep2", "C_Rep3"), "_OR8"),
                        Syn_Cond2 = paste0("t00_", c("A_Rep1", "C_Rep1", "B_Rep3"), "_OR8"))
sel_samples_4v4 <- list(Syn_Cond1 = paste0("t00_", c("B_Rep3_OR9", "B_Rep3_OR8", "C_Rep3_OR8", "A_Rep2_OR8")),
                        Syn_Cond2 = paste0("t00_", c("C_Rep2_OR8", "C_Rep2_OR9", "A_Rep2_OR9", "B_Rep2_OR8")))
sel_samples_6v6 <- list(Syn_Cond1 =c("t00_A_Rep3_OR8", "t00_C_Rep2_OR8", "t00_C_Rep3_OR9",
                                     "t00_A_Rep2_OR9", "t00_B_Rep2_OR9", "t00_B_Rep1_OR8"),
                        Syn_Cond2 =c("t00_C_Rep1_OR8", "t00_A_Rep3_OR9", "t00_A_Rep1_OR9",
                                     "t00_B_Rep2_OR8", "t00_B_Rep1_OR9", "t00_B_Rep3_OR8"))

cond1_label <- "Condition 1"
cond2_label <- "Condition 2"
neutal_label <- "Ignored"
annotation_syn_cond <- data.frame(
  `Comparison_3v3` = case_when(
    colnames(data_norm) %in% sel_samples_3v3$Syn_Cond1 ~ cond1_label,
    colnames(data_norm) %in% sel_samples_3v3$Syn_Cond2 ~ cond2_label,
    TRUE ~ neutal_label
  ),
  `Comparison_4v4` = case_when(
    colnames(data_norm) %in% sel_samples_4v4$Syn_Cond1 ~ cond1_label,
    colnames(data_norm) %in% sel_samples_4v4$Syn_Cond2 ~ cond2_label,
    TRUE ~ neutal_label
  ),
  `Comparison_6v6` = case_when(
    colnames(data_norm) %in% sel_samples_6v6$Syn_Cond1 ~ cond1_label,
    colnames(data_norm) %in% sel_samples_6v6$Syn_Cond2 ~ cond2_label,
    TRUE ~ neutal_label
  ),
  row.names = colnames(data_norm)
)

cowplot::plot_grid(pheatmap::pheatmap(data_norm, show_rownames = FALSE, 
                   cluster_rows = hclust(protein_distances), 
                   cluster_cols = hclust(sample_distances),
                   color = viridis::magma(n=100),
                   annotation_col = annotation_syn_cond, 
                   annotation_colors = list(
                     Comparison_3v3 = c(`Condition 1` = "#b2df8a", `Condition 2` = "#1f78b4", Ignored = "grey90"),
                     Comparison_4v4 = c(`Condition 1` = "#b2df8a", `Condition 2` = "#1f78b4", Ignored = "grey90"),
                     Comparison_6v6 = c(`Condition 1` = "#b2df8a", `Condition 2` = "#1f78b4", Ignored = "grey90")
                   ),
                   annotation_legend = c(TRUE, FALSE, FALSE),
                   silent=TRUE)$gtable)
ggsave("plots/de_graaf_heatmap.pdf", width=8, height=6)
```






# Performance Comparison

Helper Method

```{r}
tidy_FDR <- function(data, predicted_FDR, known_class, ticks = ppoints(100)){
  
  # use tidy eval
  predicted_FDR <- rlang::enquo(predicted_FDR)
  known_class <- rlang::enquo(known_class)

  FDR_values <- rlang::eval_tidy(predicted_FDR, data)
  known_values <- rlang::eval_tidy(known_class, data)
  
  if(! is.logical(known_values)){
    stop("values in known_class must be of type logical.")
  }
  if(! is.numeric(FDR_values)){
    stop("values in known_class must be of type logical.")
  }
  
  if(any(FDR_values > 1, na.rm=TRUE) | any(FDR_values < 0, na.rm=TRUE)){
    stop("value in predicted_FDR must be between zero and one.")
  }
  
  map_dfr(ticks, function(thres){
    tab <- table(factor(FDR_values < thres, levels = c("TRUE", "FALSE")), 
                  factor(known_values, levels = c("TRUE", "FALSE")))
    tibble(DesiredFDR = thres, TP = tab[1,1], FP=tab[1,2], 
           ObservedFDR = ifelse(TP == 0 & FP == 0, 0, FP / (TP + FP)))
  })
  
}


```



```{r}
shuffle_dataset <- function(data, frac_changed, comparison = c("3v3", "4v4", "6v6")) {
  comparison <- match.arg(comparison, c("3v3", "4v4", "6v6"))

  if(comparison == "3v3"){
    # Select 3 vs 3 columns that show no change if 0% are changed
    selected_columns <-  paste0("t00_", c("A_Rep2", "C_Rep2", "C_Rep3",    # Cond 1
                                          "A_Rep1", "C_Rep1", "B_Rep3"),   # Cond 2
                                "_OR8")
    new_experimental_design <- c(rep("SynCond_1", 3), rep("SynCond_2", 3))
  }else if(comparison == "4v4"){
    # Select 4 vs 4 columns that show no change if 0% are changed
    selected_columns <-  paste0("t00_", c("B_Rep3_OR9", "B_Rep3_OR8", "C_Rep3_OR8", "A_Rep2_OR8",    # Cond 1
                                          "C_Rep2_OR8", "C_Rep2_OR9", "A_Rep2_OR9", "B_Rep2_OR8")    # Cond 2
                                )
    new_experimental_design <- c(rep("SynCond_1", 4), rep("SynCond_2", 4))
  }else{
    selected_columns <-  c("t00_A_Rep3_OR8", "t00_C_Rep2_OR8", "t00_C_Rep3_OR9", 
                           "t00_A_Rep2_OR9", "t00_B_Rep2_OR9", "t00_B_Rep1_OR8",   # Cond1
                           "t00_C_Rep1_OR8", "t00_A_Rep3_OR9", "t00_A_Rep1_OR9", 
                           "t00_B_Rep2_OR8", "t00_B_Rep1_OR9", "t00_B_Rep3_OR8")   # Cond2
    new_experimental_design <- c(rep("SynCond_1", 6), rep("SynCond_2", 6))
  }
  
  data_changed <- data[, selected_columns]
  

  changed <- rep(FALSE, nrow(data_changed))
    changed[sample(seq_len(nrow(data_changed)), size=round(nrow(data_changed) * frac_changed))] <- TRUE
  changed_indices <- which(changed)
  
  data_changed[changed_indices, new_experimental_design == "SynCond_1"] <- 
    data_changed[sample(changed_indices), new_experimental_design == "SynCond_1"]
  
  list(
    data = data_changed,
    experimental_design = new_experimental_design,
    changed = changed
  )
}
```






```{r}
imp_methods <- list(
  list("zero"),
  list("QRILC"),
  list("knn", colmax = 0.95),
  list("MinDet"),
  list("MinProb")
)

```




Beautiful plot

```{r}
color_mapping <- c(proDA = "#5a8c08", QPROT = "#fb8072", Perseus = "#386cb0",
                   `knn-limma`="#5B127F", `QRILC-limma`="#D070FF", `MinDet-limma`="#B724FF",
                   `MinProb-limma`="#68387F", `zero-limma`="#921DCC")
background_painter <- data.frame(DesiredFDR=c(0, 0.1, 0.2), ObservedFDR=c(0, 0.1, 0.2))
```



```{r}

make_calib_and_perf_plot <- function(tp_fp_df){

  calib_plot <- tp_fp_df %>%
    ggplot(aes(x=DesiredFDR, y=ObservedFDR, group=method, color=method)) +
      geom_ribbon(data=background_painter, aes(ymax=ObservedFDR), ymin=0, color="white", fill="#91bfdb", alpha=0.1, group=99) +
      geom_ribbon(data=background_painter, aes(ymin=ObservedFDR), ymax=1, color="white", fill="#fc8d59", alpha=0.1, group=98) +
      annotate("text", x=0.085, y=0.03, hjust=0, label="conservative", size=5, alpha=0.5) +
      annotate("text", x=0.03, y=0.085, hjust=0, angle = 90, label="anti-conservative", size=5, alpha=0.5) +
      annotate("text", x=0.15, y=0.16, hjust=0, angle=45, label="optimal", alpha=0.5, size=5) +
      geom_abline(intercept=0, slope=1, size=2, color="grey") +
      geom_line() +
      geom_line(data=. %>% filter(method == "proDA"), size=1.5, color=color_mapping["proDA"]) +
      ggrepel::geom_label_repel(data=. %>% 
                     filter(DesiredFDR == max(DesiredFDR)),
                aes(label=nice_label, color= method),
                text.color = "black", segment.colour = "lightgrey", 
                anchor.side = "left", size=5,
                direction="y", hjust=1, xlim=c(0.215, NA)) +
      scale_color_manual(values= color_mapping,  guide="none") + 
      scale_x_continuous(breaks =  c(0, 0.05, 0.1, 0.15, 0.2), 
                         labels =  c("0", "0.05", "0.10", "0.15", "0.20"),
                         expand = expand_scale(add = 0.001, 0)) +
      scale_y_continuous(breaks =  c(0, 0.05, 0.1, 0.15, 0.2), 
                         labels =  c("0", "0.05", "0.10", "0.15", "0.20"),
                         expand = expand_scale(add = c(0.001, 0))) +
      lemon::coord_flex_fixed(xlim = c(0, 0.3), ylim=c(0,0.2), 
                              bottom = lemon::capped_horizontal("right", gap=0)) +
      xlab(paste0("Desired FDR", paste0(rep(" ", 32), collapse = ""))) + ylab("Actual FDR")
  
  perf_plot <- tp_fp_df %>%
    ggplot(aes(x=DesiredFDR, y=TP, group=method, color=method)) +
      geom_line() +
      geom_line(data=. %>% filter(method == "proDA"), size=1.5, color=color_mapping["proDA"]) +
      ggrepel::geom_label_repel(data=. %>% 
                     filter(DesiredFDR == max(DesiredFDR)),
                aes(label=nice_label, color= method),
                text.color = "black", segment.colour = "lightgrey", 
                anchor.side = "left", size=5,
                direction="y", hjust=1, xlim=c(0.215, NA)) +
      scale_color_manual(values= color_mapping,  guide="none") + 
      scale_x_continuous(breaks =  c(0, 0.05, 0.1, 0.15, 0.2), 
                         labels =  c("0", "0.05", "0.10", "0.15", "0.20")) +
      ylim(0, NA) +
      lemon::coord_capped_cart(xlim = c(0, 0.3), gap = 0, 
                              expand = expand_scale(add = c(0.01, 0.1)),
                              bottom='right') +
      xlab(paste0("Desired FDR", paste0(rep(" ", 32), collapse = ""))) + ylab("True Positives") 
  list(calib_plot, perf_plot)
}

```





```{r}
run_tools <- function(perc_changed, comparison, methods = c("proDA", "DEP", "QPROT", "Perseus")){
  
  if(! exists("data_norm")){
    stop("data_norm variable not yet initialized")
  }
  if("DEP" %in% methods && ! exists("imp_methods")){
    stop("imp_methods variable not yet initialized")
  }
  
  set.seed(1)
  shuf_df <- shuffle_dataset(data_norm, perc_changed, comparison = comparison)
  
  if("proDA" %in% methods){
    proDA_result <- run_proDA(shuf_df$data,
                              shuf_df$experimental_design, verbose=TRUE)
    proDA_result_df <- as_tibble(proDA_result$result) %>%
      mutate(method = "proDA",
             changed = shuf_df$changed)  %>% 
      select(name, pval, adj_pval, diff, method, changed)
  }else{
    proDA_result_df <- NULL
  }

  if("DEP" %in% methods){
    dep_result_df <- map_dfr(imp_methods, function(imp_met){
      print(paste0("Do ", imp_met[[1]]))
      tmp <- do.call(run_dep, c(list(shuf_df$data,
                                     experimental_design = shuf_df$experimental_design,
                     imputation_method = imp_met[[1]]), imp_met[-1]))
      tmp$method <- paste0(imp_met[[1]], "-limma")
      tmp$changed <- shuf_df$changed
      tmp
    }) %>%
    select(name, pval, adj_pval, diff, method, changed)
  }else{
    dep_result_df <- NULL
  }
  
  if("QPROT" %in% methods){
    qprot_result <- run_qprot(shuf_df$data[, shuf_df$experimental_design == "SynCond_1"],
                              shuf_df$data[, shuf_df$experimental_design == "SynCond_2"])
    qprot_result_df <- as_tibble(qprot_result) %>%
      transmute(name=Protein, pval = pnorm(Zstatistic), adj_pval = fdr, diff=LogFoldChange) %>%
      mutate(pval = pmin(pval, 1-pval) * 2) %>%
      mutate(method = "QPROT",
             changed = shuf_df$changed) %>%
      select(name, pval, adj_pval, diff, method, changed)
  }else{
    qprot_result_df <- NULL
  }

  if("Perseus" %in% methods){
    perseus_result<- run_perseus(shuf_df$data, tag = paste0(perc_changed * 100, "_", comparison), 
                                 return_null_if_missing = TRUE)
    if(is.null(perseus_result)){
      perseus_result_df <- NULL
    }else{
      perseus_result_df <- as_tibble(perseus_result) %>%
        mutate(method = "Perseus",
               changed = shuf_df$changed)  %>%
        select(name, pval, adj_pval, diff, method, changed) 
    }
  }else{
    perseus_result_df <- NULL
  }

  null_df <- data.frame(method = character(0),
                        name = character(0),
                        pval = numeric(0),
                        adj_pval = numeric(0),
                        diff = numeric(0),
                        changed = logical(0),
                        stringsAsFactors = FALSE)
                        
  res_df <- bind_rows(list(null_df, dep_result_df ,
    proDA_result_df,
    qprot_result_df ,
    perseus_result_df))
  
  list(shuf_df = shuf_df, res_df = res_df)
}

run_performance_comparison <- function(perc_changed, comparison,
                                       methods = c("proDA", "DEP", "QPROT", "Perseus")){
  run_tools(perc_changed, comparison, methods)$res_df %>%
    group_by(method) %>%
    do({tidy_FDR(., adj_pval, changed, ticks = ppoints(101) * 0.2)}) %>%
    mutate(nice_label = case_when(
      method == "zero-limma" ~ "DEP: Zero",
      method == "knn-limma" ~ "DEP: KNN",
      str_detect(method, "limma") ~ paste0("DEP: ", str_remove(method, "-limma")),
      TRUE ~ method
    ))
}


```







```{r}
tp_df_5_6v6 <- run_performance_comparison(perc_changed = 0.05, comparison = "6v6", methods = c("DEP", "proDA", "QPROT", "Perseus")) 
tp_df_10_6v6 <- run_performance_comparison(perc_changed = 0.1, comparison = "6v6", methods = c("DEP", "proDA", "QPROT", "Perseus")) 
tp_df_20_6v6 <- run_performance_comparison(perc_changed = 0.2, comparison = "6v6", methods = c("DEP", "proDA", "QPROT", "Perseus")) 
tp_df_30_6v6 <- run_performance_comparison(perc_changed = 0.3, comparison = "6v6", methods = c("DEP", "proDA", "QPROT", "Perseus")) 

tp_df_5_4v4 <- run_performance_comparison(perc_changed = 0.05, comparison = "4v4", methods = c("DEP", "proDA", "QPROT", "Perseus")) 
tp_df_10_4v4 <- run_performance_comparison(perc_changed = 0.1, comparison = "4v4", methods = c("DEP", "proDA", "QPROT", "Perseus")) 
tp_df_20_4v4 <- run_performance_comparison(perc_changed = 0.2, comparison = "4v4", methods = c("DEP", "proDA", "QPROT", "Perseus")) 
tp_df_30_4v4 <- run_performance_comparison(perc_changed = 0.3, comparison = "4v4", methods = c("DEP", "proDA", "QPROT", "Perseus")) 

tp_df_5_3v3 <- run_performance_comparison(perc_changed = 0.05, comparison = "3v3", methods = c("DEP", "proDA", "QPROT", "Perseus")) 
tp_df_10_3v3 <- run_performance_comparison(perc_changed = 0.1, comparison = "3v3", methods = c("DEP", "proDA", "QPROT", "Perseus")) 
tp_df_20_3v3 <- run_performance_comparison(perc_changed = 0.2, comparison = "3v3", methods = c("DEP", "proDA", "QPROT", "Perseus")) 
tp_df_30_3v3 <- run_performance_comparison(perc_changed = 0.3, comparison = "3v3", methods = c("DEP", "proDA", "QPROT", "Perseus")) 
```



```{r}
saveRDS(list(
  tp_df_5_6v6=tp_df_5_6v6,tp_df_10_6v6=tp_df_10_6v6,tp_df_20_6v6=tp_df_20_6v6,tp_df_30_6v6=tp_df_30_6v6,
  tp_df_5_4v4=tp_df_5_4v4,tp_df_10_4v4=tp_df_10_4v4,tp_df_20_4v4=tp_df_20_4v4,tp_df_30_4v4=tp_df_30_4v4,
  tp_df_5_3v3=tp_df_5_3v3,tp_df_10_3v3=tp_df_10_3v3,tp_df_20_3v3=tp_df_20_3v3,tp_df_30_3v3=tp_df_30_3v3),
  "tmp/de_graaf/3v3_4v4_6v6-results.RDS")
```


```{r}

all_tp_df <- readRDS("tmp/de_graaf/3v3_4v4_6v6-results.RDS")

tp_df_5_6v6 <- rbind(all_tp_df$tp_df_5_6v6)
tp_df_5_4v4 <- rbind(all_tp_df$tp_df_5_4v4)
tp_df_5_3v3 <- rbind(all_tp_df$tp_df_5_3v3)
tp_df_10_6v6 <- rbind(all_tp_df$tp_df_10_6v6)
tp_df_10_4v4 <- rbind(all_tp_df$tp_df_10_4v4)
tp_df_10_3v3 <- rbind(all_tp_df$tp_df_10_3v3)
tp_df_20_6v6 <- rbind(all_tp_df$tp_df_20_6v6)
tp_df_20_4v4 <- rbind(all_tp_df$tp_df_20_4v4)
tp_df_20_3v3 <- rbind(all_tp_df$tp_df_20_3v3)
tp_df_30_6v6 <- rbind(all_tp_df$tp_df_30_6v6)
tp_df_30_4v4 <- rbind(all_tp_df$tp_df_30_4v4)
tp_df_30_3v3 <- rbind(all_tp_df$tp_df_30_3v3)
```



```{r}

all_plots <- lapply(list(tp_df_5_6v6, tp_df_10_6v6, tp_df_20_6v6, tp_df_30_6v6,
                         tp_df_5_4v4, tp_df_10_4v4, tp_df_20_4v4, tp_df_30_4v4,
                         tp_df_5_3v3, tp_df_10_3v3, tp_df_20_3v3, tp_df_30_3v3), 
                    make_calib_and_perf_plot)

```



```{r}
annotation_box <- function(perc){
  ggdraw() +
    draw_grob(grid::rectGrob(just="top", y=1, height=0.85, width=0.5 ,gp=grid::gpar(fill="grey80", lty=0))) +
    draw_text(paste0(perc, "% changed"), y=0.6, angle = 90) 
}

```



```{r}
set.seed(1)
pl_6v6 <- plot_grid(ggdraw() + draw_text("6 vs 6 samples", size=20, fontface = "bold"),
    plot_grid(
      plot_grid(annotation_box(5),
          plot_grid(plotlist = all_plots[[1]], labels=c("A", "B")),
      rel_widths = c(1, 10), nrow=1),
      plot_grid(annotation_box(10),
          plot_grid(plotlist = all_plots[[2]], labels=c("C", "D")),
      rel_widths = c(1, 10), nrow=1),
      plot_grid(annotation_box(20),
          plot_grid(plotlist = all_plots[[3]], labels=c("E", "F")),
      rel_widths = c(1, 10), nrow=1),
      plot_grid(annotation_box(30),
          plot_grid(plotlist = all_plots[[4]], labels=c("G", "H")),
      rel_widths = c(1, 10), nrow=1), ncol=1),
rel_heights = c(1, 20), ncol=1)
ggplot2::ggsave("plots/calib_perf_plot-6v6.pdf", pl_6v6, width = 14, height=16)

pl_4v4 <- plot_grid(ggdraw() + draw_text("4 vs 4 samples", size=20, fontface = "bold"),
    plot_grid(
      plot_grid(annotation_box(5),
          plot_grid(plotlist = all_plots[[5]], labels=c("A", "B")),
      rel_widths = c(1, 10), nrow=1),
      plot_grid(annotation_box(10),
          plot_grid(plotlist = all_plots[[6]], labels=c("C", "D")),
      rel_widths = c(1, 10), nrow=1),
      plot_grid(annotation_box(20),
          plot_grid(plotlist = all_plots[[7]], labels=c("E", "F")),
      rel_widths = c(1, 10), nrow=1),
      plot_grid(annotation_box(30),
          plot_grid(plotlist = all_plots[[8]], labels=c("G", "H")),
      rel_widths = c(1, 10), nrow=1), ncol=1),
rel_heights = c(1, 20), ncol=1)
ggplot2::ggsave("plots/calib_perf_plot-4v4.pdf", pl_4v4, width = 14, height=16)


pl_3v3 <- plot_grid(ggdraw() + draw_text("3 vs 3 samples", size=20, fontface = "bold"),
    plot_grid(
      plot_grid(annotation_box(5),
          plot_grid(plotlist = all_plots[[9]], labels=c("A", "B")),
      rel_widths = c(1, 10), nrow=1),
      plot_grid(annotation_box(10),
          plot_grid(plotlist = all_plots[[10]], labels=c("C", "D")),
      rel_widths = c(1, 10), nrow=1),
      plot_grid(annotation_box(20),
          plot_grid(plotlist = all_plots[[11]], labels=c("E", "F")),
      rel_widths = c(1, 10), nrow=1),
      plot_grid(annotation_box(30),
          plot_grid(plotlist = all_plots[[12]], labels=c("G", "H")),
      rel_widths = c(1, 10), nrow=1), ncol=1),
rel_heights = c(1, 20), ncol=1)
ggplot2::ggsave("plots/calib_perf_plot-3v3.pdf", pl_3v3, width = 14, height=16)
```




```{r}
plot_grid(plotlist = all_plots[[11]], labels=c("A", "B"), nrow=1)
ggsave("plots/calib_perf_plot_main_text.pdf", width=12, height=4)
```







Why do the tools differ?

```{r}

tmp_res <- run_tools(perc_changed = 0.2, comparison = "3v3", methods = c("DEP", "proDA", "QPROT", "Perseus")) 

```

```{r}
write_rds(tmp_res, "tmp/de_graaf/3v3_20_changed_raw_results.RDS")
tmp_res <- read_rds( "tmp/de_graaf/3v3_20_changed_raw_results.RDS")
```


```{r}
tmp_res$res_df %>%
 group_by(method) %>%
    do({tidy_FDR(., adj_pval, changed, ticks = ppoints(101) * 0.2)}) %>%
    mutate(nice_label = case_when(
      method == "zero-limma" ~ "DEP: Zero",
      method == "knn-limma" ~ "DEP: KNN",
      str_detect(method, "limma") ~ paste0("DEP: ", str_remove(method, "-limma")),
      TRUE ~ method
    )) %>%
  make_calib_and_perf_plot()
```


```{r}
obs_df <- tibble(name = rownames(tmp_res$shuf_df$data),
                 Obs1 = rowSums(! is.na(tmp_res$shuf_df$data[, 1:3])),
                 Obs2 = rowSums(! is.na(tmp_res$shuf_df$data[, 4:6]))) %>%
  mutate(ObsLabel = paste0(pmin(Obs1, Obs2), "-", pmax(Obs1, Obs2))) %>%
  mutate(ObsLabel = fct_reorder2( factor(ObsLabel, levels = c("All-All", unique(ObsLabel))), -pmin(Obs1, Obs2), -pmax(Obs1, Obs2)))

tp_fp_per_obs_10 <- tmp_res$res_df %>%
  inner_join(obs_df) %>%
  mutate(adj_pval = ifelse(is.na(adj_pval), 1, adj_pval)) %>%
  group_by(method, ObsLabel) %>%
  summarize(TP = sum(changed & adj_pval < 0.1),
            FP = sum(!changed & adj_pval < 0.1),
            MargT = sum(changed)) %>%
  bind_rows(tmp_res$res_df %>%
    distinct() %>%
    inner_join(obs_df) %>%
    mutate(adj_pval = ifelse(is.na(adj_pval), 1, adj_pval)) %>%
    group_by(method) %>%
    summarize(TP = sum(changed & adj_pval < 0.1),
              FP = sum(!changed & adj_pval < 0.1),
              MargT = sum(changed)) %>%
    mutate(ObsLabel = "All-All")) %>%
  mutate(FDR = ifelse(FP == 0, 0, FP / (FP + TP))) %>%
  mutate(nice_label = case_when(
      method == "zero-limma" ~ "DEP: Zero",
      method == "knn-limma" ~ "DEP: KNN",
      str_detect(method, "limma") ~ paste0("DEP: ", str_remove(method, "-limma")),
      TRUE ~ method
    ))


tp_fp_per_obs_01 <- tmp_res$res_df %>%
  inner_join(obs_df) %>%
  mutate(adj_pval = ifelse(is.na(adj_pval), 1, adj_pval)) %>%
  group_by(method, ObsLabel) %>%
  summarize(TP = sum(changed & adj_pval < 0.01),
            FP = sum(!changed & adj_pval < 0.01),
            MargT = sum(changed)) %>%
  bind_rows(tmp_res$res_df %>%
    distinct() %>%
    inner_join(obs_df) %>%
    mutate(adj_pval = ifelse(is.na(adj_pval), 1, adj_pval)) %>%
    group_by(method) %>%
    summarize(TP = sum(changed & adj_pval < 0.01),
              FP = sum(!changed & adj_pval < 0.01),
              MargT = sum(changed)) %>%
    mutate(ObsLabel = "All-All")) %>%
  mutate(FDR = ifelse(FP == 0, 0, FP / (FP + TP))) %>%
  mutate(nice_label = case_when(
      method == "zero-limma" ~ "DEP: Zero",
      method == "knn-limma" ~ "DEP: KNN",
      str_detect(method, "limma") ~ paste0("DEP: ", str_remove(method, "-limma")),
      TRUE ~ method
    ))


fdr_per_obs_plot_10 <- ggplot(tp_fp_per_obs_10, aes(x = nice_label, y = ObsLabel)) +
    geom_tile(aes(fill = FDR)) +
    geom_text(aes(label  = paste0(FP, " / ", (FP + TP))), size=1.8) +
    geom_hline(yintercept = 10.53, color="black") +
    theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1)) +
    scale_fill_gradient2(low = "#91bfdb", mid="white", high = "#fc8d59",
                         midpoint = 0.1,
                         limits = c(0, 0.2), 
                         expand = expand_scale(add = c(0, 0)),
                         oob = function(x, range = c(0,1), only.finite){
                           pmax(range[1], pmin(range[2], x))
                         }) +
  ggtitle("FDR by Observations", "(FP / Considered Positive)")  +
  ylab("Observations per Condition") + xlab("")

tp_per_obs_plot_10 <- ggplot(tp_fp_per_obs_10, aes(x = nice_label, y = ObsLabel)) +
  geom_tile(data = . %>% filter(ObsLabel != "All-All"), aes(fill = TP)) +
  geom_tile(data = . %>% filter(ObsLabel == "All-All"), fill="white") +
  geom_text(aes(label  = paste0(TP, " / ", MargT)), size=1.8) +
  geom_hline(yintercept = 10.53, color="black") +
  theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1)) +
  scale_fill_gradientn(colors = scales::brewer_pal(palette = "Purples", direction = -1)(9)[3:9],
                       name = "TP") +
  ggtitle("TP by Observations", "(TP / Actually Positive)") +
  ylab("Observations per Condition") + xlab("")


fdr_per_obs_plot_01 <- ggplot(tp_fp_per_obs_01, aes(x = nice_label, y = ObsLabel)) +
    geom_tile(aes(fill = FDR)) +
    geom_text(aes(label  = paste0(FP, " / ", (FP + TP))), size=1.8) +
    geom_hline(yintercept = 10.53, color="black") +
    theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1)) +
    scale_fill_gradient2(low = "#91bfdb", mid="white", high = "#fc8d59",
                         midpoint = 0.01,
                         limits = c(0, 0.02), 
                         expand = expand_scale(add = c(0, 0)),
                         oob = function(x, range = c(0,1), only.finite){
                           pmax(range[1], pmin(range[2], x))
                         }) +
  ggtitle("FDR by Observations", "(FP / Considered Positive)")  +
  ylab("Observations per Condition") + xlab("")



tp_per_obs_plot_01 <- ggplot(tp_fp_per_obs_01, aes(x = nice_label, y = ObsLabel)) +
  geom_tile(data = . %>% filter(ObsLabel != "All-All"), aes(fill = TP)) +
  geom_tile(data = . %>% filter(ObsLabel == "All-All"), fill="white") +
  geom_text(aes(label  = paste0(TP, " / ", MargT)), size=1.8) +
  geom_hline(yintercept = 10.53, color="black") +
  theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1)) +
  scale_fill_gradientn(colors = scales::brewer_pal(palette = "Purples", direction = -1)(9)[3:9],
                       name = "TP") +
  ggtitle("TP by Observations", "(TP / Actually Positive)") +
  ylab("Observations per Condition") + xlab("")


fdr_10_box <- ggdraw() +
    draw_grob(grid::rectGrob(just="top", y=1, height=0.9, width=0.5 ,gp=grid::gpar(fill="grey80", lty=0))) +
    draw_text(paste0("Desired FDR fixed at 10%"), y=0.55, angle = 90) 

fdr_01_box <- ggdraw() +
    draw_grob(grid::rectGrob(just="top", y=1, height=0.9, width=0.5 ,gp=grid::gpar(fill="grey80", lty=0))) +
    draw_text(paste0("Desired FDR fixed at  1%"), y=0.55, angle = 90) 

plot_grid(fdr_10_box, fdr_per_obs_plot_10, tp_per_obs_plot_10,
          fdr_01_box, fdr_per_obs_plot_01, tp_per_obs_plot_01, 
          nrow=2, labels=c("", "A", "B", "", "C", "D"),
          rel_widths = c(1, 10, 10))

ggsave("plots/performance_by_observation.pdf", width=10, height=8)

```








```{r}


 ggplot(tp_fp_per_obs_10, aes(x = nice_label, y = ObsLabel)) +
    geom_tile(data = . %>% filter(ObsLabel != "All-All"), aes(fill = TP)) +
    geom_tile(data = . %>% filter(ObsLabel == "All-All"), fill="white") +
    geom_text(aes(label  = paste0(TP, " / ", MargT)), size=2) +
    theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1)) +
    scale_fill_gradientn(colors = scales::brewer_pal(palette = "Purples", direction = -1)(9)[4:9],
                         name = "TP") +
    ggtitle("TP by Observations", "(TP / Actually Positive)") +
    ylab("Observations per Condition") + xlab("")
```






# Session Info

```{r}
sessionInfo()
```






