---
title: "R Notebook"
output: html_notebook
---






```{r}

library(tidyverse)
library(proDA)
library(DEP)
library(DAPAR)
library(SummarizedExperiment)
library(cowplot)
source("run_dep.R")
source("run_proDA.R")
source("run_qprot.R")
source("run_perseus.R")
source("run_dapar.R")
source("run_ebrct.R")
source("RanCenHier2Fun_modified.R")

```


Load the data from the de Graaf et al. (2014) pubication that studied the effect of 
prostaglandin E2 (PGE2) on phosphorylation at different timepoints. The data was
uploaded to the ProteomeXchange with the identifier PXD000293. I have extracted
the proteinGroups MaxQuant output file. I will now load that file

```{r}

full_data <- read_tsv(
  "../data/degraaf_phosphorylation/degraaf_proteinGroups.txt",
  col_types = list(
    Reverse = "c"
  )
)

tidy_data <- full_data %>%
  filter(is.na(Reverse) & is.na(`Contaminant`)) %>%
  dplyr::select(ProteinID=`Protein IDs`, starts_with("LFQ Intensity ")) %>%
  gather(Sample, Intensity, starts_with("LFQ intensity ")) %>%
  mutate(MS_Run = str_match(Sample, "^LFQ intensity (OR\\d)_(\\d+)min_([A-C])(\\d)$")[,2]) %>% 
  mutate(Timepoint = as.numeric(str_match(Sample, "^LFQ intensity (OR\\d)_(\\d+)min_([A-C])(\\d)$")[,3])) %>% 
  mutate(BioReplicate = str_match(Sample, "^LFQ intensity (OR\\d)_(\\d+)min_([A-C])(\\d)$")[,4]) %>% 
  mutate(TechReplicate = as.numeric(str_match(Sample, "^LFQ intensity (OR\\d)_(\\d+)min_([A-C])(\\d)$")[,5])) %>% 
  mutate(SampleName = paste0("t", str_pad(Timepoint, width=2, pad="0"),
                             "_", BioReplicate, "_Rep", TechReplicate,  "_", MS_Run)) 

tidy_data

```




```{r}

data <- tidy_data %>%
  filter(Timepoint == 0) %>%
  mutate(Intensity = ifelse(Intensity == 0, NA, log2(Intensity))) %>%
  select(ProteinID, SampleName, Intensity) %>%
  spread(SampleName, Intensity) %>%
  column_to_rownames("ProteinID") %>%
  as.matrix() 

data <- data[rowSums(is.na(data)) < 18, ]

data_norm <- median_normalization(data)



annotation_df <- tidy_data %>%
  filter(Timepoint == 0) %>%
  select(SampleName, BioReplicate, TechReplicate, MS_Run) %>%
  distinct() %>%
  mutate(TechReplicate = as.factor(TechReplicate)) %>%
  arrange(BioReplicate, TechReplicate) %>%
  as.data.frame() %>%
  column_to_rownames("SampleName")


```



```{r}
data_for_plot <- data_norm
data_for_plot[is.na(data_for_plot)] <- 0
pheatmap::pheatmap(data_for_plot, show_rownames = FALSE, 
                   annotation_col = annotation_df)

```



```{r}
set.seed(1)
zero_parameters <- proDA(data_norm, col_data = annotation_df, verbose=TRUE)

```

```{r}
# write_rds(zero_parameters, "tmp/de_graaf/zero_info_parameters.RDS")
# zero_parameters <- read_rds("tmp/de_graaf/zero_info_parameters.RDS")
```



```{r}
sample_distances <- dist_approx(zero_parameters)$mean
protein_distances <- dist_approx( zero_parameters, by_sample = FALSE)$mean
```

```{r}
cowplot::plot_grid(pheatmap::pheatmap(data_norm, show_rownames = FALSE, 
                   cluster_rows = hclust(protein_distances), 
                   cluster_cols = hclust(sample_distances),
                   annotation_col = annotation_df, silent=TRUE)$gtable)


as.data.frame(cmdscale(sample_distances)) %>%
  rownames_to_column("Names") %>%
  inner_join(rownames_to_column(annotation_df, "Names")) %>%
  ggplot(aes(x = V1, y=V2, color=MS_Run, shape = BioReplicate)) +
    geom_point(size=4) +
    ggrepel::geom_text_repel(aes(label = Names), point.padding = 0.1, box.padding = 0.3, size=3) +
    coord_fixed()
```


```{r}
annotation_df <- tidy_data %>%
  filter(Timepoint == 0) %>%
  select(SampleName, BioReplicate, TechReplicate, MS_Run) %>%
  distinct() %>%
  mutate(TechReplicate = as.factor(TechReplicate)) %>%
  arrange(BioReplicate, TechReplicate) %>%
  as.data.frame() %>%
  column_to_rownames("SampleName")


sel_samples_3v3 <- list(Syn_Cond1 = paste0("t00_", c("A_Rep2", "C_Rep2", "C_Rep3"), "_OR8"),
                        Syn_Cond2 = paste0("t00_", c("A_Rep1", "C_Rep1", "B_Rep3"), "_OR8"))
sel_samples_4v4 <- list(Syn_Cond1 = paste0("t00_", c("B_Rep3_OR9", "B_Rep3_OR8", "C_Rep3_OR8", "A_Rep2_OR8")),
                        Syn_Cond2 = paste0("t00_", c("C_Rep2_OR8", "C_Rep2_OR9", "A_Rep2_OR9", "B_Rep2_OR8")))
sel_samples_6v6 <- list(Syn_Cond1 =c("t00_A_Rep3_OR8", "t00_C_Rep2_OR8", "t00_C_Rep3_OR9",
                                     "t00_A_Rep2_OR9", "t00_B_Rep2_OR9", "t00_B_Rep1_OR8"),
                        Syn_Cond2 =c("t00_C_Rep1_OR8", "t00_A_Rep3_OR9", "t00_A_Rep1_OR9",
                                     "t00_B_Rep2_OR8", "t00_B_Rep1_OR9", "t00_B_Rep3_OR8"))

cond1_label <- "Condition 1"
cond2_label <- "Condition 2"
neutal_label <- "Ignored"
annotation_syn_cond <- data.frame(
  `Comparison_3v3` = case_when(
    colnames(data_norm) %in% sel_samples_3v3$Syn_Cond1 ~ cond1_label,
    colnames(data_norm) %in% sel_samples_3v3$Syn_Cond2 ~ cond2_label,
    TRUE ~ neutal_label
  ),
  `Comparison_4v4` = case_when(
    colnames(data_norm) %in% sel_samples_4v4$Syn_Cond1 ~ cond1_label,
    colnames(data_norm) %in% sel_samples_4v4$Syn_Cond2 ~ cond2_label,
    TRUE ~ neutal_label
  ),
  `Comparison_6v6` = case_when(
    colnames(data_norm) %in% sel_samples_6v6$Syn_Cond1 ~ cond1_label,
    colnames(data_norm) %in% sel_samples_6v6$Syn_Cond2 ~ cond2_label,
    TRUE ~ neutal_label
  ),
  row.names = colnames(data_norm)
)

cowplot::plot_grid(pheatmap::pheatmap(data_norm, show_rownames = FALSE, 
                   cluster_rows = hclust(protein_distances), 
                   cluster_cols = hclust(sample_distances),
                   color = viridis::magma(n=100),
                   annotation_col = annotation_syn_cond, 
                   annotation_colors = list(
                     Comparison_3v3 = c(`Condition 1` = "#b2df8a", `Condition 2` = "#1f78b4", Ignored = "grey90"),
                     Comparison_4v4 = c(`Condition 1` = "#b2df8a", `Condition 2` = "#1f78b4", Ignored = "grey90"),
                     Comparison_6v6 = c(`Condition 1` = "#b2df8a", `Condition 2` = "#1f78b4", Ignored = "grey90")
                   ),
                   annotation_legend = c(TRUE, FALSE, FALSE),
                   silent=TRUE)$gtable)
# ggsave("plots/de_graaf_heatmap.pdf", width=8, height=6)
```






# Performance Comparison

Helper Method

```{r}
tidy_FDR <- function(data, predicted_FDR, known_class, ticks = ppoints(100)){
  
  # use tidy eval
  predicted_FDR <- rlang::enquo(predicted_FDR)
  known_class <- rlang::enquo(known_class)

  FDR_values <- rlang::eval_tidy(predicted_FDR, data)
  known_values <- rlang::eval_tidy(known_class, data)
  
  if(! is.logical(known_values)){
    stop("values in known_class must be of type logical.")
  }
  if(! is.numeric(FDR_values)){
    stop("values in known_class must be of type logical.")
  }
  
  if(any(FDR_values > 1, na.rm=TRUE) | any(FDR_values < 0, na.rm=TRUE)){
    stop("value in predicted_FDR must be between zero and one.")
  }
  
  map_dfr(ticks, function(thres){
    tab <- table(factor(FDR_values < thres, levels = c("TRUE", "FALSE")), 
                  factor(known_values, levels = c("TRUE", "FALSE")))
    tibble(DesiredFDR = thres, TP = tab[1,1], FP=tab[1,2], 
           ObservedFDR = ifelse(TP == 0 & FP == 0, 0, FP / (TP + FP)))
  })
  
}


```



```{r}
shuffle_dataset <- function(data, frac_changed, comparison = c("3v3", "4v4", "6v6")) {
  comparison <- match.arg(comparison, c("3v3", "4v4", "6v6"))

  if(comparison == "3v3"){
    # Select 3 vs 3 columns that show no change if 0% are changed
    selected_columns <-  paste0("t00_", c("A_Rep2", "C_Rep2", "C_Rep3",    # Cond 1
                                          "A_Rep1", "C_Rep1", "B_Rep3"),   # Cond 2
                                "_OR8")
    new_experimental_design <- c(rep("SynCond_1", 3), rep("SynCond_2", 3))
  }else if(comparison == "4v4"){
    # Select 4 vs 4 columns that show no change if 0% are changed
    selected_columns <-  paste0("t00_", c("B_Rep3_OR9", "B_Rep3_OR8", "C_Rep3_OR8", "A_Rep2_OR8",    # Cond 1
                                          "C_Rep2_OR8", "C_Rep2_OR9", "A_Rep2_OR9", "B_Rep2_OR8")    # Cond 2
                                )
    new_experimental_design <- c(rep("SynCond_1", 4), rep("SynCond_2", 4))
  }else{
    selected_columns <-  c("t00_A_Rep3_OR8", "t00_C_Rep2_OR8", "t00_C_Rep3_OR9", 
                           "t00_A_Rep2_OR9", "t00_B_Rep2_OR9", "t00_B_Rep1_OR8",   # Cond1
                           "t00_C_Rep1_OR8", "t00_A_Rep3_OR9", "t00_A_Rep1_OR9", 
                           "t00_B_Rep2_OR8", "t00_B_Rep1_OR9", "t00_B_Rep3_OR8")   # Cond2
    new_experimental_design <- c(rep("SynCond_1", 6), rep("SynCond_2", 6))
  }
  
  data_changed <- data[, selected_columns]
  

  changed <- rep(FALSE, nrow(data_changed))
    changed[sample(seq_len(nrow(data_changed)), size=round(nrow(data_changed) * frac_changed))] <- TRUE
  changed_indices <- which(changed)
  
  data_changed[changed_indices, new_experimental_design == "SynCond_1"] <- 
    data_changed[sample(changed_indices), new_experimental_design == "SynCond_1"]
  
  list(
    data = data_changed,
    experimental_design = new_experimental_design,
    changed = changed
  )
}
```






```{r}
imp_methods <- list(
  list("zero"),
  list("QRILC"),
  list("knn", colmax = 0.95),
  list("MinDet"),
  list("MinProb")
)

```




Beautiful plot

```{r}
color_mapping <- c(proDA = "#33A02C", 
                   `knn-limma`="#CAB2D6", `QRILC-limma`="#B294C7", `MinDet-limma`="#9A77B8", `MinProb-limma`="#825AA9", `zero-limma`="#6A3D9A",
                   DAPAR = "#1F78B4", 
                   Perseus = "#8AADBF",
                   QPROT = "#bfa419", EBRCT = "#B15928")
background_painter <- data.frame(DesiredFDR=c(0, 0.1, 0.2), ObservedFDR=c(0, 0.1, 0.2))

plot(c(0, 1), c(0, 1), type="n")
rect(seq(0, 1-1/length(color_mapping), l=length(color_mapping)), 0, seq(1/length(color_mapping), 1, length=length(color_mapping)), 1, col=color_mapping)
text(x=seq(0, 1-1/length(color_mapping), l=length(color_mapping)) + 0.5/length(color_mapping), y=0.7, labels=color_mapping, srt=90)
text(x=seq(0, 1-1/length(color_mapping), l=length(color_mapping)) + 0.5/length(color_mapping), y=0.3, labels=names(color_mapping), srt=90)

```






```{r}
run_tools <- function(perc_changed, comparison, methods = c("proDA", "DEP", "QPROT", "Perseus", "DAPAR", "EBRCT")){
  
  if(! exists("data_norm")){
    stop("data_norm variable not yet initialized")
  }
  if("DEP" %in% methods && ! exists("imp_methods")){
    stop("imp_methods variable not yet initialized")
  }
  
  set.seed(1)
  shuf_df <- shuffle_dataset(data_norm, perc_changed, comparison = comparison)
  changed_df <- tibble(name = rownames(data_norm), changed = shuf_df$changed)
  whole_row_missing <- rowSums(is.na(shuf_df$data)) == ncol(shuf_df$data)
  shuf_df$data <- shuf_df$data[! whole_row_missing, ]
  changed_df <- changed_df[! whole_row_missing, ]
  
  if("proDA" %in% methods){
    proDA_result <- run_proDA(shuf_df$data,
                              shuf_df$experimental_design, verbose=TRUE)
    proDA_result_df <- as_tibble(proDA_result$result) %>%
      mutate(method = "proDA") %>% 
      inner_join(changed_df) %>%
      select(name, pval, adj_pval, diff, method, changed)
  }else{
    proDA_result_df <- NULL
  }
  
  if("EBRCT" %in% methods){
     ebrct_res <- run_ebrct(shuf_df$data, 
                            shuf_df$experimental_design)
     ebrct_res_df <-  ebrct_res %>%
       mutate(method = "EBRCT") %>% 
       inner_join(changed_df) %>%
       select(name, pval, adj_pval, diff, method, changed)
   }else{
      ebrct_res_df <- NULL
   }
  
  if("DAPAR" %in% methods){
    dapar_result <- run_dapar(shuf_df$data,
                              shuf_df$experimental_design)
    dapar_result_df <- as_tibble(dapar_result) %>%
      mutate(method = "DAPAR") %>% 
      inner_join(changed_df) %>%
      select(name, pval, adj_pval, diff, method, changed)
  }else{
    dapar_result_df <- NULL
  }

  if("DEP" %in% methods){
    dep_result_df <- map_dfr(imp_methods, function(imp_met){
      print(paste0("Do ", imp_met[[1]]))
      tmp <- do.call(run_dep, c(list(shuf_df$data,
                                     experimental_design = shuf_df$experimental_design,
                     imputation_method = imp_met[[1]]), imp_met[-1]))
      tmp$method <- paste0(imp_met[[1]], "-limma")
      tmp <- inner_join(tmp, changed_df)
      tmp
    }) %>%
    select(name, pval, adj_pval, diff, method, changed)
  }else{
    dep_result_df <- NULL
  }
  
  if("QPROT" %in% methods){
    qprot_result <- run_qprot(shuf_df$data[, shuf_df$experimental_design == "SynCond_1"],
                              shuf_df$data[, shuf_df$experimental_design == "SynCond_2"])
    qprot_result_df <- as_tibble(qprot_result) %>%
      transmute(name=Protein, pval = pnorm(Zstatistic), adj_pval = fdr, diff=-LogFoldChange) %>%
      mutate(pval = pmin(pval, 1-pval) * 2) %>%
      mutate(method = "QPROT") %>% 
      inner_join(changed_df) %>%
      select(name, pval, adj_pval, diff, method, changed)
  }else{
    qprot_result_df <- NULL
  }

  if("Perseus" %in% methods){
    perseus_result<- run_perseus(shuf_df$data, tag = paste0(perc_changed * 100, "_", comparison),
                                 return_null_if_missing = TRUE,
                                 dataset_name = "de_graaf/null_comparison")
    if(is.null(perseus_result)){
      perseus_result_df <- NULL
    }else{
      perseus_result_df <- as_tibble(perseus_result) %>%
        mutate(method = "Perseus") %>% 
        inner_join(changed_df) %>%
        select(name, pval, adj_pval, diff, method, changed) 
    }
  }else{
    perseus_result_df <- NULL
  }

  null_df <- data.frame(method = character(0),
                        name = character(0),
                        pval = numeric(0),
                        adj_pval = numeric(0),
                        diff = numeric(0),
                        changed = logical(0),
                        stringsAsFactors = FALSE)
                        
  res_df <- bind_rows(list(null_df, dep_result_df ,
    proDA_result_df,
    qprot_result_df ,
    perseus_result_df,
    dapar_result_df,
    ebrct_res_df))
  
  list(shuf_df = shuf_df, res_df = res_df)
}


```



```{r}
res_3v3 <- run_tools(0, "3v3", c("proDA", "Perseus", "QPROT", "EBRCT", "DAPAR", "DEP"))$res_df
res_4v4 <- run_tools(0, "4v4", c("proDA", "Perseus", "QPROT", "EBRCT", "DAPAR", "DEP"))$res_df
res_6v6 <- run_tools(0, "6v6", c("proDA", "Perseus", "QPROT", "EBRCT", "DAPAR", "DEP"))$res_df
```


Load Triqler Results:


```{r}

load_triqler_results <- function(comparison){
  comparison <- match.arg(comparison, c("3v3", "4v4", "6v6"))
  
  tr_list <- read_lines(paste0("tmp/de_graaf/triqler_results/triqler_result_", comparison, ".tsv")) %>%
    str_split("\t")
  tr_headers <- tr_list[[1]]
  
  triqler_result_df_raw <- tr_list[-1] %>%
    map_dfr(function(chr){
      c(set_names(as.list(chr[which(tr_headers != "peptides")]), tr_headers[tr_headers != "peptides"]),
        list(peptides = paste0(chr[-which(tr_headers != "peptides")], collapse=";")))
    }) %>%
    transmute(protein = protein, peptides = peptides,
              q_value = parse_number(q_value),
              posterior_error_prob = parse_number(posterior_error_prob),
              num_peptides = parse_number(num_peptides),
              log2_fold_change = parse_number(log2_fold_change),
              diff_exp_prob_0.0 = parse_number(diff_exp_prob_0.0))
  
  triqler_result_df_raw %>%
    mutate(Origin = case_when(
      str_starts(protein, "REV__") ~ "Decoy",
      str_starts(protein, "CON__") ~ "Contamination",
      TRUE ~ "Other"
    )) %>%
    filter(Origin != "Decoy" & Origin != "Contamination")  %>%
    transmute(name = protein, pval = NA, adj_pval = q_value,
                diff = -log2_fold_change) %>%
    mutate(method = "Triqler",
           changed = FALSE)

}


res_3v3 <- rbind(res_3v3, load_triqler_results("3v3"))
res_4v4 <- rbind(res_4v4, load_triqler_results("4v4"))
res_6v6 <- rbind(res_6v6, load_triqler_results("6v6"))

```


```{r}
res_3v3 <- res_3v3 %>%
    mutate(nice_label = case_when(
      method == "zero-limma" ~ "DEP: Zero",
      method == "knn-limma" ~ "DEP: KNN",
      str_detect(method, "limma") ~ paste0("DEP: ", str_remove(method, "-limma")),
      TRUE ~ method
    ))

res_4v4 <- res_4v4 %>%
    mutate(nice_label = case_when(
      method == "zero-limma" ~ "DEP: Zero",
      method == "knn-limma" ~ "DEP: KNN",
      str_detect(method, "limma") ~ paste0("DEP: ", str_remove(method, "-limma")),
      TRUE ~ method
    ))

res_6v6 <- res_6v6 %>%
    mutate(nice_label = case_when(
      method == "zero-limma" ~ "DEP: Zero",
      method == "knn-limma" ~ "DEP: KNN",
      str_detect(method, "limma") ~ paste0("DEP: ", str_remove(method, "-limma")),
      TRUE ~ method
    ))


```





```{r}

make_false_positive_plots <- function(res){

  hist_plot <- res %>%
    filter(method != "Triqler") %>%
    mutate(signif_detect = adj_pval > 0.1) %>%
    ggplot(aes(x = pval, fill = signif_detect)) +
      geom_histogram(breaks = seq(0, 1, length.out = 21), color = "white", size=0) +
      scale_x_continuous(breaks = c(0, 0.5, 1), labels = c("0", "0.5", "1"), name = "p-value histogram") +
      scale_y_continuous(breaks = seq(0, 1000, by = 50), labels = rep("", 21), name = "", expand = expand_scale(add = c(0, 10))) + 
      # scale_fill_viridis_d(direction = -1) +
      scale_fill_manual(values = c("FALSE"= "#dea11d", "TRUE" = "#410373")) +
      facet_wrap(~ nice_label, scales = "free_y", nrow=1) +
      theme(legend.position = "none", axis.text.y = element_blank(), strip.text = element_text(size = 10))
  
  
  dlog_scatter <- res %>%
    filter(method != "Triqler") %>%
    group_by(method) %>%
    arrange(pval) %>%
    mutate(uniform_distr = ppoints(n())) %>%
    mutate(signif_detect = factor(adj_pval < 0.1, levels = c(TRUE, FALSE),
                                  labels = c("adjusted p-value < 0.1", "adjusted p-value > 0.1"),
                                  ordered = TRUE)) %>%
    ggplot(aes(x = uniform_distr, y = pval)) +
      geom_point(data = . %>% filter( adj_pval < 0.1), aes(color = signif_detect)) + 
      geom_point(data = . %>% filter( adj_pval >= 0.1), aes(color = signif_detect)) + 
      geom_abline() +
      scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
                    labels = c("", "0.01", "", "1"),
                    minor_breaks = c(0.001, 0.005, 0.01, 0.05, 0.1, 0.5, 1)) + 
      scale_y_log10(breaks = c(0.001, 0.01, 0.1, 1),
                    labels = c("", "0.01", "", "1"),
                    minor_breaks =  c(0.001, 0.005, 0.01, 0.05, 0.1, 0.5, 1)) +
      # scale_color_viridis_d(direction = -1) + 
      scale_color_manual(values = c("adjusted p-value < 0.1"= "#dea11d", "adjusted p-value > 0.1" = "#410373")) +
      ylab("Sorted p-values") + xlab("Uniform values between 0 and 1") +
      facet_wrap(~ nice_label, nrow=1) +
      guides(colour = guide_legend(override.aes = list(size = 5), title = "")) + 
      theme(legend.position = "bottom", 
            panel.grid.minor = element_line(color = alpha("grey", 0.2), size = 0.3),
            panel.grid.major = element_line(color = alpha("grey", 0.2), size = 0.3), 
            strip.text = element_text(size = 10))
  
  
  tab <- res %>%
    group_by(nice_label) %>%
    summarize(FP = sum(adj_pval < 0.1, na.rm=TRUE)) %>%
    arrange(FP)
  
  tab_grob <- tab %>%
    rowwise() %>%
    mutate(FP = format(FP, width = 1)) %>%
    gridExtra::tableGrob(cols = c("Method", "False\nPositives"), rows = NULL,
                                   theme = gridExtra::ttheme_minimal(core = list(fg_params = list(hjust = 0, x = 0.05, vjust = 0.5)),
                                                                     colhead = list(fg_params = list(vjust = 0, y = 0.2, hjust = 0, x = 0.05))))
  
  
  cowplot::plot_grid(
    tab_grob,
    cowplot::plot_grid(hist_plot, NULL, dlog_scatter, nrow=3, align = "vh", labels = c("B", ""), rel_heights = c(0.7, 0.2, 1)),
    nrow = 1, rel_widths = c(1, 5), labels = c("A", "")
  )
}


```


```{r}
make_false_positive_plots(res_3v3)
ggsave("plots/de_graaf_null_comparison_3v3.pdf", width = 13, height = 5)

make_false_positive_plots(res_4v4)
ggsave("plots/de_graaf_null_comparison_4v4.pdf", width = 13, height = 5)

make_false_positive_plots(res_6v6)
ggsave("plots/de_graaf_null_comparison_6v6.pdf", width = 13, height = 5)


```





```{r}

# library(MSqRob)
# library(MSnbase)
# 
# mse <- MSqRob::import2MSnSet("../data/degraaf_null_comparison/peptides.txt",
#                              filetype = "MaxQuant")
# 
# 
# exprs(mse) <- ifelse(exprs(mse) == 0, NA, log2(exprs(mse)))
#   
# 
# mse_norm <- proDA::median_normalization(mse)
# 
# selected_columns <-  paste0("OR8_0min_", c("A2", "C2", "C3",    # Cond 1
#                                            "A1", "C1", "B3"),   # Cond 2
#                             "")
# experimental_design <- c(rep("SynCond_1", 3), rep("SynCond_2", 3))
# 
# exp_annotation <- data.frame(run = selected_columns, experimental_design=experimental_design,
#                              row.names = selected_columns, stringsAsFactors = FALSE)
# mse_norm_sub <- mse_norm[, selected_columns]
# 
# pData(mse_norm_sub) <- exp_annotation
# 
# proteins_mse_norm <- MSnSet2protdata(mse_norm_sub, accession = "Proteins",
#                                      printProgress = TRUE)
# # Filter out empty proteins
# proteins_mse_norm_filtered <- proteins_mse_norm[which(sapply(seq_len(length(proteins_mse_norm)), function(x) nrow(getData(proteins_mse_norm[x]))) != 0)]
# 
# prot_fit <- fit.model(proteins_mse_norm_filtered, response = "quant_value",
#                       fixed = "experimental_design", random = c("run", "Sequence"),
#                       add.intercept = TRUE, weights="Huber")
# 
# cntrst <- makeContrast("experimental_designSynCond_2 - experimental_designSynCond_1", 
#                        c("experimental_designSynCond_2", "experimental_designSynCond_1"))
# prot_result <- test.protLMcontrast(prot_fit, cntrst)
# prot_result_annotated <- prot_result %>% 
#   as_tibble(rownames = "name") %>%
#   mutate(adj_pval = p.adjust(pval, "BH")) %>%
#   mutate(diff = estimate) %>%
#   mutate(id = seq_len(n())) %>%
#   mutate(Protein = str_split(name, ";")) %>%
#   unnest() %>%
#   transmute(name, pval, adj_pval, diff)

```



```{r}
# hist(prot_result_annotated$pval)
# table(is.na(prot_result_annotated$pval))
# plot(ppoints(sum(! is.na(prot_result_annotated$pval))), sort(prot_result_annotated$pval), log="xy"); abline(0,1)
# 
# summary(prot_result_annotated$adj_pval < 0.1)

```




```{r}
# library(msEmpiRe)
# 
# # Make experimental design file
# # tribble(
# #   ~Name, ~Condition,
# #   "H1",   "H",
# #   "H2",   "H",
# #   "H3",   "H",
# #   "L1",   "L",
# #   "L2",   "L",
# #   "L3",   "L",
# # ) %>%
# tibble(Name = selected_columns, Condition = experimental_design) %>%
#   write_tsv("/tmp/condition_sample_mapping.tsv")
# 
# # Load Data
# msemp <- read.MaxQuant("../data/degraaf_null_comparison/peptides.txt", "/tmp/condition_sample_mapping.tsv")
# 
# # Filter out stuff based on the proteinGroups.txt file
# msemp <- filter_MaxQuant(msemp, "../data/cox_proteome_benchmark/proteinGroups.txt", filter_on = c("Contaminant", "Reverse"))
# 
# # Filter out unused samples
# msemp <- msemp[, ! is.na(msemp$condition)]
# 
# # do log transformation
# msemp_cp <- msemp
# exprs(msemp_cp) <- ifelse(exprs(msemp_cp) == 0, NA, log2(exprs(msemp_cp)))
# 
# # Normalize data
# msemp_norm3 <- proDA::median_normalization(msemp_cp)
# 
# # Transform back to un-logarithmized scale
# msemp_cp2 <- msemp_norm3
# 
# exprs(msemp_cp2) <- ifelse(is.na(exprs(msemp_cp2)), 0, 2^(exprs(msemp_cp2)))
# # Filter out proteins that are completely missing in one condition
# cond_compl_mis <- rowSums(exprs(msemp_cp2)[,pData(msemp_cp2)$condition == "SynCond_1"] == 0) == 3 |
#   rowSums(exprs(msemp_cp2)[,pData(msemp_cp2)$condition == "SynCond_2"] == 0) == 3
# 
# # Do the actual test
# dir.create("/tmp/msempire_plots")
# msemp_res <- de.ana(msemp_cp2[! cond_compl_mis, ], out.dir = "/tmp/msempire_plots/")
# 
# 
# msemp_res_annotated <- as_tibble(msemp_res) %>%
#   transmute(name = prot.id, pval = p.val, adj_pval = p.adj, diff=-log2FC) 
```


```{r}
hist(msemp_res_annotated$pval)
table(msemp_res_annotated$adj_pval < 0.1, useNA = "always")
plot(ppoints(sum(! is.na(msemp_res_annotated$pval))), sort(msemp_res_annotated$pval), log="xy"); abline(0,1)
```















##----------------Legacy-----------------##



```{r}
saveRDS(list(
),
  "tmp/null_comparison/3v3_4v4_6v6-results.RDS")
```


```{r}

all_tp_df <- readRDS("tmp/null_comparison/3v3_4v4_6v6-results.RDS")


```










```{r}
tmp_res$res_df %>%
 group_by(method) %>%
    do({tidy_FDR(., adj_pval, changed, ticks = ppoints(101) * 0.2)}) %>%
    mutate(nice_label = case_when(
      method == "zero-limma" ~ "DEP: Zero",
      method == "knn-limma" ~ "DEP: KNN",
      str_detect(method, "limma") ~ paste0("DEP: ", str_remove(method, "-limma")),
      TRUE ~ method
    )) %>%
  make_calib_and_perf_plot()
```





```{r}
obs_df <- tibble(name = rownames(tmp_res$shuf_df$data),
                 Obs1 = rowSums(! is.na(tmp_res$shuf_df$data[, 1:3])),
                 Obs2 = rowSums(! is.na(tmp_res$shuf_df$data[, 4:6]))) %>%
  mutate(ObsLabel = paste0(pmin(Obs1, Obs2), "-", pmax(Obs1, Obs2))) %>%
  mutate(ObsLabel = fct_reorder2( factor(ObsLabel, levels = c("All-All", unique(ObsLabel))), -pmin(Obs1, Obs2), -pmax(Obs1, Obs2)))

ref_df <- tibble(name = rownames(tmp_res$shuf_df$data), changed = tmp_res$shuf_df$changed) %>%
  inner_join(obs_df) %>%
  group_by(ObsLabel) %>%
  summarize(TP = 0,
            FP = 0,
            MargT = sum(changed)) 


tp_fp_per_obs_10 <- tmp_res$res_df %>%
  inner_join(obs_df) %>%
  mutate(adj_pval = ifelse(is.na(adj_pval), 1, adj_pval)) %>%
  group_by(method, ObsLabel) %>%
  summarize(TP = sum(changed & adj_pval < 0.1),
            FP = sum(!changed & adj_pval < 0.1),
            MargT = sum(changed)) %>%
  group_by(method) %>%
  group_map(function(df, key){
    bind_rows(df, ref_df %>% filter(! ObsLabel %in% df$ObsLabel))
  }) %>%
  bind_rows(tmp_res$res_df %>%
    distinct() %>%
    inner_join(obs_df) %>%
    mutate(adj_pval = ifelse(is.na(adj_pval), 1, adj_pval)) %>%
    group_by(method) %>%
    summarize(TP = sum(changed & adj_pval < 0.1),
              FP = sum(!changed & adj_pval < 0.1),
              MargT = sum(changed)) %>%
    mutate(ObsLabel = "All-All")) %>%
  mutate(FDR = ifelse(FP == 0, 0, FP / (FP + TP))) %>%
  mutate(nice_label = case_when(
      method == "zero-limma" ~ "DEP: Zero",
      method == "knn-limma" ~ "DEP: KNN",
      str_detect(method, "limma") ~ paste0("DEP: ", str_remove(method, "-limma")),
      TRUE ~ method
    ))


tp_fp_per_obs_01 <- tmp_res$res_df %>%
  inner_join(obs_df) %>%
  mutate(adj_pval = ifelse(is.na(adj_pval), 1, adj_pval)) %>%
  group_by(method, ObsLabel) %>%
  summarize(TP = sum(changed & adj_pval < 0.01),
            FP = sum(!changed & adj_pval < 0.01),
            MargT = sum(changed)) %>%
  group_by(method) %>%
  group_map(function(df, key){
    bind_rows(df, ref_df %>% filter(! ObsLabel %in% df$ObsLabel))
  }) %>%
  bind_rows(tmp_res$res_df %>%
    distinct() %>%
    inner_join(obs_df) %>%
    mutate(adj_pval = ifelse(is.na(adj_pval), 1, adj_pval)) %>%
    group_by(method) %>%
    summarize(TP = sum(changed & adj_pval < 0.01),
              FP = sum(!changed & adj_pval < 0.01),
              MargT = sum(changed)) %>%
    mutate(ObsLabel = "All-All")) %>%
  mutate(FDR = ifelse(FP == 0, 0, FP / (FP + TP))) %>%
  mutate(nice_label = case_when(
      method == "zero-limma" ~ "DEP: Zero",
      method == "knn-limma" ~ "DEP: KNN",
      str_detect(method, "limma") ~ paste0("DEP: ", str_remove(method, "-limma")),
      TRUE ~ method
    ))


fdr_per_obs_plot_10 <- ggplot(tp_fp_per_obs_10, aes(x = nice_label, y = ObsLabel)) +
    geom_tile(aes(fill = FDR)) +
    geom_text(aes(label  = paste0(FP, " / ", (FP + TP))), size=1.8) +
    geom_hline(yintercept = 10.53, color="black") +
    theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1)) +
    scale_fill_gradient2(low = "#91bfdb", mid="white", high = "#fc8d59",
                         midpoint = 0.1,
                         limits = c(0, 0.2), 
                         expand = expand_scale(add = c(0, 0)),
                         oob = function(x, range = c(0,1), only.finite){
                           pmax(range[1], pmin(range[2], x))
                         }) +
  ggtitle("FDR by Observations", "(FP / Considered Positive)")  +
  ylab("Observations per Condition") + xlab("")

tp_per_obs_plot_10 <- ggplot(tp_fp_per_obs_10, aes(x = nice_label, y = ObsLabel)) +
  geom_tile(data = . %>% filter(ObsLabel != "All-All"), aes(fill = TP)) +
  geom_tile(data = . %>% filter(ObsLabel == "All-All"), fill="white") +
  geom_text(aes(label  = paste0(TP, " / ", MargT)), size=1.8) +
  geom_hline(yintercept = 10.53, color="black") +
  theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1)) +
  scale_fill_gradientn(colors = scales::brewer_pal(palette = "Purples", direction = -1)(9)[3:9],
                       name = "TP") +
  ggtitle("TP by Observations", "(TP / Actually Positive)") +
  ylab("Observations per Condition") + xlab("")


fdr_per_obs_plot_01 <- ggplot(tp_fp_per_obs_01, aes(x = nice_label, y = ObsLabel)) +
    geom_tile(aes(fill = FDR)) +
    geom_text(aes(label  = paste0(FP, " / ", (FP + TP))), size=1.8) +
    geom_hline(yintercept = 10.53, color="black") +
    theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1)) +
    scale_fill_gradient2(low = "#91bfdb", mid="white", high = "#fc8d59",
                         midpoint = 0.01,
                         limits = c(0, 0.02), 
                         expand = expand_scale(add = c(0, 0)),
                         oob = function(x, range = c(0,1), only.finite){
                           pmax(range[1], pmin(range[2], x))
                         }) +
  ggtitle("FDR by Observations", "(FP / Considered Positive)")  +
  ylab("Observations per Condition") + xlab("")



tp_per_obs_plot_01 <- ggplot(tp_fp_per_obs_01, aes(x = nice_label, y = ObsLabel)) +
  geom_tile(data = . %>% filter(ObsLabel != "All-All"), aes(fill = TP)) +
  geom_tile(data = . %>% filter(ObsLabel == "All-All"), fill="white") +
  geom_text(aes(label  = paste0(TP, " / ", MargT)), size=1.8) +
  geom_hline(yintercept = 10.53, color="black") +
  theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1)) +
  scale_fill_gradientn(colors = scales::brewer_pal(palette = "Purples", direction = -1)(9)[3:9],
                       name = "TP") +
  ggtitle("TP by Observations", "(TP / Actually Positive)") +
  ylab("Observations per Condition") + xlab("")


fdr_10_box <- ggdraw() +
    draw_grob(grid::rectGrob(just="top", y=1, height=0.9, width=0.5 ,gp=grid::gpar(fill="grey80", lty=0))) +
    draw_text(paste0("Desired FDR fixed at 10%"), y=0.55, angle = 90) 

fdr_01_box <- ggdraw() +
    draw_grob(grid::rectGrob(just="top", y=1, height=0.9, width=0.5 ,gp=grid::gpar(fill="grey80", lty=0))) +
    draw_text(paste0("Desired FDR fixed at  1%"), y=0.55, angle = 90) 

plot_grid(fdr_10_box, fdr_per_obs_plot_10, tp_per_obs_plot_10,
          fdr_01_box, fdr_per_obs_plot_01, tp_per_obs_plot_01, 
          nrow=2, labels=c("", "A", "B", "", "C", "D"),
          rel_widths = c(1, 10, 10))

# ggsave("plots/performance_by_observation.pdf", width=10, height=8)

```








```{r}


 ggplot(tp_fp_per_obs_10, aes(x = nice_label, y = ObsLabel)) +
    geom_tile(data = . %>% filter(ObsLabel != "All-All"), aes(fill = TP)) +
    geom_tile(data = . %>% filter(ObsLabel == "All-All"), fill="white") +
    geom_text(aes(label  = paste0(TP, " / ", MargT)), size=2) +
    theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1)) +
    scale_fill_gradientn(colors = scales::brewer_pal(palette = "Purples", direction = -1)(9)[4:9],
                         name = "TP") +
    ggtitle("TP by Observations", "(TP / Actually Positive)") +
    ylab("Observations per Condition") + xlab("")
```






# Session Info

```{r}
sessionInfo()
```







